On shift equivalence for non-linear rules\thanks{Research partially supported by the MIUR PRIN 2017FTXR7S ``IT-MaTTerS''.}

### 2023-07-16 Related/GraphShift.tex
some related stuff

%We extend Soft Concurrent Constraint languages with the possibility to manage both local (and global) knowledge of agents. Being constraints defined as soft, the underlying system of preferences consists in an ordered monoid where it is possible to natively represent partially ordered and bipolar (i.e., positive/negative) preferences. Such a formalism further generalise the previous literature, while it also grants  a \emph{polynomial}  representation of soft constraints that can be exploited in  several applications. Variables, which support constraints, can either be private to a single agent, or in common with all of them, thus allowing  for different levels of visibility. We provide labelled and unlabelled reduction semantics and their correspondence, together with a labelled bisimulation equivalence.

@@

The fact that if we switch two steps and then switch them back, we may obtain a derivation that is not coincident with the original one has unpleasant consequences. For instance, if we restrict $\rho_1'$ and $\rho_3'$ to the first two steps, we get non-equivalent derivations $\delta_1;\delta_2 \not\shifteq \delta_1;\delta_2'$, but $\rho_1 \shifteq[\sigma] \rho_3$, with $\sigma$ mapping $\delta_1$ to $\delta_1$ and $\delta_2$ to $\delta_2'$.
%
Such a behaviour with respect to composition does not allow to obtain a causal semantics for graph transformation systems.
More precisely, it does  not allow to associate an event structure to a derivation.
%
Thus, the notion of sequential independence has to be refined. The point is that in the associated shift equivalence we want to forget who made the fusion, 
but we should not forget (as it happens in the example) that some fusions occurred.

@@

\begin{proof}
  To be done. The fact that sequential independence coincides with
  independence for contiguous derivations is easy.
\end{proof}

@@

  Moreover, let $\upsilon_1, \upsilon_2$ be adjacent
  transpositions and $\sigma' = \upsilon_1; \ldots, \upsilon_m$ a
  minimal length decomposition of $\sigma'$ in terms of adjacent
  transpositions (that thus operate on inversions). Then, for all
  $h \in \interval{m}$ with $\upsilon_h = \transp{l}[m]$ there exists
  $k \in \interval{m}$ such that
  $\upsilon_k = \transp{\upsilon_{1,h-1}^{-1};v_1;
    \upsilon_{1,k-1}(l)}[\upsilon_{1,h-1}^{-1};v_1; \upsilon_{1,k-1}(m)]$.
\end{lemma}

@@

\bibliographystyle{splncs04}
%\bibliographystyle{plainurl}
\bibliography{Unstable}

>>>>

@@

The fact that if we switch two steps and then switch them back, we obtain a derivation that is not equivalent to the original one has unpleasant consequences. For instance, if we restrict $\rho_1'$ and $\rho_3'$ to the first two steps, we get non-equivalent derivations $\delta_1;\delta_2 \not\shifteq \delta_1;\delta_2'$, but $\rho_1 \shifteq[\sigma] \rho_3$, with $\sigma$ mapping $\delta_1$ to $\delta_1$ and $\delta_2$ to $\delta_2'$.
%
The lack of such closure with respect to composition does not allow to obtain a causal semantics for graph transformation systems.
More precisely, it does  not allow to associate an event structure to a derivation.
%
Thus, the notion of sequential independence has to be refined. The point is that in the associated shift equivalence we want to forget who made the fusion, 
but we should not forget (as it happens in the example) that some fusions occurred.

@@

\begin{proof}
  To be done. The fact that sequential independence coincides with
  independence for contiguous derivations is easy. The fact that it coincides when derivations are not contiguous is false (the approach based on defining independence for non-contiguous derivation is not working, but the results should be true ...)
\end{proof}

@@

  Moreover, let $\upsilon_1, \upsilon_2$ be adjacent
  transpositions and $\sigma' = \upsilon_1; \ldots, \upsilon_m$ a
  minimal length decomposition of $\sigma'$ in terms of adjacent
  transpositions (that thus operate on inversions). Then, for all
  $h \in \interval{m}$ with $\upsilon_h = \transp{l}[l+1]$ there exists
  $k \in \interval{m}$ such that
  $\upsilon_k = \transp{\upsilon_{1,h-1}^{-1};v_1;
    \upsilon_{1,k-1}(l)}[\upsilon_{1,h-1}^{-1};v_1; \upsilon_{1,k-1}(l+1)]$.
\end{lemma}

@@
